A Test of the Frequency Independence Assumption of the System Parameters used in Geomagnetically Induced Current Calculations

### 2019-01-11 agujournaltemplate.tex
Updates from ShareLaTeX

%% March 2018
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def, etc.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
% DO NOT USE \psfrag or \subfigure commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%

%% To submit your paper:
\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.

@@

% Now, type in the journal name: \journalname{<Journal Name>}

% ie, \journalname{Journal of Geophysical Research}
%% Choose from this list of Journals:
%
% JGR-Atmospheres
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Oceans
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%

@@

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{=enter title here=}

@@

\authors{=list all authors here=}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

@@

\correspondingauthor{=name=}{=email address=}

%% Keypoints, final entry on title page.

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation
% \end{keypoints}

@@

\begin{keypoints}
\item = enter point 1 here =
\item = enter point 2 here =
\item = enter point 3 here =
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

@@

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

@@

Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth’s surface.  These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less and the movement of Earth’s surface relative to slow-varying current systems in Earth’s ionosphere that are near stationary relative to the Sun [REF].  Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure [REF].  Accurate estimation of GICs given E or B is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.

A common assumption made in estimates of GIC using either measured or estimated values of E is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.  The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods [e.g., Boteler, 2014a; Boteler and Pirjola, 2014b]. This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012}.

@@

One-second-cadence geoelectric field and geomagnetic field data were obtained from the Kakioka Magnetic Observatory of the Japan Meteorological Agency data portal [http://www.kakioka-jma.go.jp/obsdata/metadata/en]. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure 1.

@@

\acknowledgments
 = enter acknowledgments here =

>>>>

\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.

@@

@@

\begin{document}
\title{A Test of the Frequency Independence Assumption of the System Parameters used in Geomagnetically Induced Current Calculations}

@@

\authors{R.S. Weigel\affil{1} and P.J. Cilliers\affil{2}}

\affiliation{1}{Space Weather Lab, George Mason University}
\affiliation{2}{South African National Space Agency}

\affiliation{1}{4400 University Drive, Fairfax VA 22030}
\affiliation{2}{Hospital Street, Hermanus 7200}

%(repeat as many times as is necessary)

@@

\correspondingauthor{R.S. Weigel}{rweigel@gmu.edu}

@@

\begin{keypoints}
\item The system parameters derived from GIC and geoelectric field measurements are frequency dependent
\item A model with frequency dependence significantly outperforms a frequency independent model in predicting GIC
\item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field
\end{keypoints}

@@

@@

Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth’s surface.  These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less and the movement of Earth’s surface relative to slow-varying current systems in Earth’s ionosphere that are near stationary relative to the Sun [REF].  Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.  Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.

A common assumption made in estimates of GIC using either measured or estimated values of E is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.  The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods [e.g., Boteler, 2014a; Boteler and Pirjola, 2014b]. This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.

@@

One-second-cadence geoelectric field and geomagnetic field data were obtained from the Kakioka Magnetic Observatory of the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure 1.

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal http://www.kakioka-jma.go.jp/obsdata/metadata/en.
